---
title: "Bacteria Tufts"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Bacteria Tufts}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

## Description

Basic details about the assessment. Update the 'response' values as required.

```{r description}

description <- tribble(
  ~question, ~response,
  "name_short", "Bacterial Tufts",
  "name_long", "Bacterial Tufts Metric",
  "parameter", "SEWAGE_FUNGUS",
  "status", "prototype",
  "type", "metric"
)

description
```

## Input

A list of questions required to run the assessment.

```{r input}

input <- tibble(
  location_id = c("8175", "8175"),
  sample_id = c("12345", "12345"),
  date_taken = as.Date(c("2019-11-21", "2019-11-21")),
  question = c("Sewage fungus cover", "Sewage fungus density"),
  response = c("<30%", "Thin"),
  label = c(NA, NA),
  parameter = c("SEWAGE_FUNGUS", "SEWAGE_FUNGUS"),
  type = c("character", "character"),
  max = c(NA, NA),
  min = c(NA, NA),
  source = c("sepa_ecology_results","sepa_ecology_results")
)
input
```

## Assessment

If applicable, write a function to assess your input data and return an outcome. For example, a metric, statistic, prediction etc.

```{r assessment}

assessment_function <- function(data) {
  # Calculated some statistic...
  # Note, any non-standard base R library must be call using require().
  require(dplyr)
  require(tidyr)
  require(magrittr)
  require(tibble)
  require(purrr)

  data <- purrr::map_df(split(data, data$sample_id), function(sample) {
    if (nrow(sample) < 2) {
      return(NULL)
    } else {
      sample <- dplyr::arrange(sample, question)
      sample$combined <- paste(sample$response, collapse = " ")
      row <- sample[1, ] # grab a row to use as template for tufts classification
      row <- dplyr::mutate(row,
        "question" = "Bacterial Tuft Class",
        "parameter" = "Bacterial Tuft"
      )
      row <- dplyr::mutate(row, "response" = dplyr::case_when(
        combined == "<30% Trace" ~ "Moderate",
        combined == "30-60% Trace" ~ "Moderate",
        combined == ">60% Trace" ~ "Poor",
        combined == "<30% Thin" ~ "Moderate",
        combined == "30-60% Thin" ~ "Poor",
        combined == ">60% Thin" ~ "Poor",
        combined == "<30% Thick" ~ "Poor",
        combined == "30-60% Thick" ~ "Poor",
        combined == ">60% Thick" ~ "Poor",
        combined == "<30% Massive" ~ "Poor",
        combined == "30-60% Massive" ~ "Bad",
        combined == ">60% Massive" ~ "Bad"
      ))

      sample <- dplyr::bind_rows(sample, row)
      sample$combined <- NULL
      return(sample)
    }
  })
  
  
  return(data)
}
```

## Outcome

The outcome of your assessment.

```{r output}
outcome <- assessment_function(input)
outcome
```

## Check

Run checks on the assessment.

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description) 
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = outcome
)
```

After **updating the catalogue, rebuild the package**, click on Build > Install and Restart menu or 'Install and Restart' button in the Build pane.

## Test

This section tests if this assessment is usable using `assessment` function.

```{r}
# No need to edit this code
assess(
  data = hera::demo_data,
  name = description$response[description$question == "name_short"]
)
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
