---
title: "Bankside Consistency"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Bankside Consistency}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

This is the Rapid Assessment Technique (RAT) for rsults from bankside invertebrates analysis to check consistency with Water Body Status.

## Description

Basic details about the assessment. Update the 'response' values as required.

```{r description}

description <- tribble(
  ~question, ~response,
  "name_short", "Bankside Consistency",
  "name_long", "Rapid Assessment of Bankside Consistency with Water Body Status",
  "parameter", "River Family Inverts",
  "status", "prototype",
  "type", "metric"
)

description
```

## Input

A list of questions required to run the assessment.

```{r input}

input <- tibble(
  location_id = c("206972"),
  sample_id = c("12345"),
  date_taken = as.Date(c("2019-11-21")),
  question = c("Taxon abundance"),
  response = c("12"),
  label = c("Baetidae"),
  parameter = c("River Family Inverts"),
  type = c("number"),
  max = NA,
  min = NA,
  source = c("sepa_ecology_results")
)
input
```

## Assessment

If applicable, write a function to assess your input data and return an outcome. For example, a metric, statistic, prediction etc.

```{r assessment}

assessment_function <- function(data) {
  # Some calculated a statistic...
  # Note, any non-standard base R library must be call using require().
  require(dplyr)
  require(tidyr)
  require(magrittr)
  require(tibble)
  # require(whpt)
  # require(macroinvertebrateMetrics)
  data$date_taken <- as.character(format.Date(data$date_taken, "%Y/%m/%d"))
    data <- filter(data, parameter == "River Family Inverts")
  metric_function <- catalogue[catalogue$assessment ==
    "Macroinvertebrate Metrics", 3][[1]]
  output <- metric_function[[1]](data)
  output <- filter(output, question %in% c("WHPT_ASPT", "WHPT_NTAXA"))
  predictors <- utils::read.csv(system.file("extdat",
    "predictors.csv",
    package = "whpt"
  ), check.names = FALSE)
  predictors$location_id <- as.character(predictors$location_id)
  predict_data <- filter(predictors, location_id %in% unique(data$location_id))
  output_location <- inner_join(output,
    data[, c(
      "location_id",
      "sample_id",
      "date_taken"
    )],
    by = "sample_id"
  )
  output_location$location_id <-
    as.character(output_location$location_id)
  whpt_input <- inner_join(output_location,
    predict_data,
    by = "location_id"
  )
  whpt_input$question[whpt_input$question == "WHPT_ASPT"] <-
    "WHPT ASPT Abund"
  whpt_input$question[whpt_input$question == "WHPT_NTAXA"] <-
    "WHPT NTAXA Abund"
  if (nrow(whpt_input) < 1) {
    return(NULL)
  } else {
    whpt_input <- unique(whpt_input)
    whpt_input$response <- as.numeric(whpt_input$response)
    consistency_check <- whpt::whpts(whpt_input)
    consistency_check$response <-
      as.character(consistency_check$response)
  }
  return(consistency_check)
}
```

## Outcome

The outcome of your assessment.

```{r output}
outcome <- assessment_function(input)
outcome
```

## Check

Run checks on the assessment.

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description)
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = outcome
)
```

After **updating the catalogue, rebuild the package**, click on Build \> Install and Restart menu or 'Install and Restart' button in the Build pane.

## Test

This section tests if this assessment is usable using `assessment` function.

```{r}
# No need to edit this code
assess(
  data = hera::demo_data,
  name = description$response[description$question == "name_short"]
)

lims_data <- hera:::convert(lims_data)
assess(
  data = lims_data,
  name = description$response[description$question == "name_short"]
)
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
