---
title: "Bankside Consistency"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Bankside Consistency}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE,
  warning = FALSE
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

This is the Rapid Assessment Technique (RAT) for results from bankside invertebrates analysis to check consistency with Water Body Status.

## Description

Basic details about the assessment.

```{r description}
description <- tribble(
  ~question, ~response,
  "name_short", "Bankside Consistency",
  "name_long", "Rapid Assessment of Bankside Consistency with Water Body Status",
  "parameter", "River Family Inverts",
  "status", "testing",
  "type", "metric"
)

knitr::kable(description)
```

## Input

A list of questions required to run the assessment.

```{r input}
input <- tibble(
  location_id = c("206972", "206972"),
  sample_id = c("12345", "54321"),
  date_taken = as.Date(c("2019-11-21", "2025-11-21")),
  question = c("Taxon abundance", "Live abundance"),
  response = c("12", "21"),
  label = c("Baetidae", "Baetidae"),
  parameter = c("River Family Inverts", "BANKSIDE_INVERTS"),
  type = c("number", "number"),
  max = c(NA, NA),
  min = c(NA, NA),
  source = c("sepa_ecology_results", "sepa_ecology_results")
)
# Only display columns vital to run example (min, max, source, type are not
# required)
input_display <- select(
  input,
  location_id,
  sample_id,
  date_taken,
  question,
  response,
  label,
  parameter
)
# Only display one example of each question
input_display <- input_display[!duplicated(input_display[, c("question")]), ]
knitr::kable(input_display)
```

## Assessment

Function code used to run this assessment.

<details>
  <summary>Code</summary>

```{r, echo = TRUE, eval = TRUE}
assessment_function <- function(data, ...) {
  # Some calculated a statistic...
  # Note, any non-standard base R library must be called using require().
  require(dplyr)
  require(tidyr)
  require(magrittr)
  require(tibble)
  require(whpt)
  require(macroinvertebrateMetrics)
  data$date_taken <- as.character(format.Date(data$date_taken, "%Y/%m/%d"))
  catalogue <- hera::catalogue
  metric_function <- catalogue[catalogue$assessment ==
    "Macroinvertebrate Metrics", 3][[1]]
  output <- metric_function[[1]](data)
  output <- filter(output, question %in% c("WHPT_ASPT", "WHPT_NTAXA"))
  predictors <- utils::read.csv(system.file("extdat",
    "gis-predictors.csv",
    package = "whpt"
  ), check.names = FALSE)

# Downgrade Typical class for testing worst-case scenario
#   predictors$`Typical ASPT Class`[predictors$`Typical ASPT Class` == "Poor"] <- "Bad"
#     predictors$`Typical ASPT Class`[predictors$`Typical ASPT Class` == "Moderate"] <- "Poor"
#       predictors$`Typical ASPT Class`[predictors$`Typical ASPT Class` == "Good"] <- "Moderate"
#         predictors$`Typical ASPT Class`[predictors$`Typical ASPT Class` == "Good"] <- "Moderate"
#           predictors$`Typical ASPT Class`[predictors$`Typical ASPT Class` == "High"] <- "Good"
# 
# 
# predictors$`Typical NTAXA Class`[predictors$`Typical NTAXA Class` == "Poor"] <- "Bad"
#     predictors$`Typical NTAXA Class`[predictors$`Typical NTAXA Class` == "Moderate"] <- "Poor"
#       predictors$`Typical NTAXA Class`[predictors$`Typical NTAXA Class` == "Good"] <- "Moderate"
#         predictors$`Typical NTAXA Class`[predictors$`Typical NTAXA Class` == "Good"] <- "Moderate"
#           predictors$`Typical NTAXA Class`[predictors$`Typical NTAXA Class` == "High"] <- "Good"

  
  predictors$location_id <- as.character(predictors$location_id)
  predict_data <- filter(predictors, location_id %in% unique(data$location_id))
  output_location <- inner_join(output,
    data[, c(
      "location_id",
      "sample_id",
      "date_taken"
    )],
    by = "sample_id",
    relationship = "many-to-many"
  )
  output_location$location_id <-
    as.character(output_location$location_id)
  whpt_input <- inner_join(output_location,
    predict_data,
    by = "location_id"
  )
  whpt_input$question[whpt_input$question == "WHPT_ASPT"] <-
    "WHPT ASPT Abund"
  whpt_input$question[whpt_input$question == "WHPT_NTAXA"] <-
    "WHPT NTAXA Abund"
  if (nrow(whpt_input) < 1) {
    return(NULL)
  } else {
    whpt_input <- unique(whpt_input)
    whpt_input$response <- as.numeric(whpt_input$response)
    consistency_check <- whpt::whpts(whpt_input)
    consistency_check$response <-
      as.character(consistency_check$response)
  }

report <- tidyr::pivot_wider(consistency_check, names_from = question, values_from = response)
vars <- c("location_id", "location_description", "sample_id", "date_taken")
location_ids <- dplyr::select(data, any_of(vars)) %>%  unique()
location_ids$season <- hera:::season(location_ids$date_taken, output = "shortname")
report <- inner_join(report, location_ids, by = join_by(sample_id))

new_predictors <- read.csv(
  system.file("extdat", "gis-predictors.csv", package = "whpt"),
  check.names = FALSE)

report <- dplyr::inner_join(report, new_predictors, by = join_by(location_id))

whpt_wide <- tidyr::pivot_wider(output, names_from = question, values_from = response)
report <- dplyr::inner_join(report, whpt_wide, by = join_by(sample_id))

vars <- c(
"water body sampled",
"sample_id", 
"date_taken",
"location_id",
"location_description",
"season",
"Reference NTAXA",
"Reference ASPT", 
"assessment",
"driver", 
"WHPT_NTAXA",
"WHPT_ASPT",
"Typical ASPT Class",
"Typical NTAXA Class",
"Reported WHPT Class Year"
)

report <- dplyr::select(report, any_of(vars))
vars <- c(
"season",
"Reference NTAXA",
"Reference ASPT", 
"assessment",
"driver", 
"WHPT_NTAXA",
"WHPT_ASPT",
"Typical ASPT Class",
"Typical NTAXA Class",
"Reported WHPT Class Year",
"water body sampled"
)
report$`water body sampled` <- as.character(report$`water body sampled`)
report$`Reported WHPT Class Year` <- as.character(report$`Reported WHPT Class Year`)
consistency_check <- pivot_longer(report, cols = all_of(vars), names_to = "question", values_to = "response")
 consistency_check$date_taken <- as.Date(consistency_check$date_taken)
  consistency_check$parameter <- "Bankside Consistency"
  return(consistency_check)
}
```

</details>

## Outcome

The outcome of your assessment.

```{r output, message=FALSE}
outcome <- assessment_function(input)
outcome_table <- select(outcome, question, response)
outcome_table <- outcome_table[!duplicated(outcome_table[, c("question")]), ]
knitr::kable(outcome_table)
```

## Check

Run checks on the assessment.

```{r checklist, message=FALSE}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description)
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update, message=FALSE}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = outcome
)
```

After **updating the catalogue, rebuild the package**, click on Build \> Install and Restart menu or 'Install and Restart' button in the Build pane.


## Test

This section tests if this assessment is usable using `assess()` function.

```{r test}
# No need to edit this code
# assess(
#   data = hera::demo_data,
#   name = description$response[description$question == "name_short"]
# )
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
