---
title: "Combined aquatic plants"
pkgdown:
  as_is: yes
output: 
  rmarkdown::html_vignette:
    toc: yes
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Combined aquatic plants}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F
)
library(hera)
library(tidyverse)
library(testthat)
```

## Welcome

This document has been created following the generic assessment guidance.


## Details

<!-- Create a new regulatory assessment by changing the CAPITLIZE info below: -->

```{r setup}

standard <- tibble(
  standard_short = NA,
  quality_element = NA,
  parameter = "Combined aquatic plants",
  standard_long = NA,
  standard_reference = NA,
  aggregation = NA,
  status = NA
)

standard_format <- hera:::hera_format(standard = standard) # format assessment data
standard_format$standard %>%
  filter(attribute != "standard_reference") %>%
  knitr::kable()
```

## Data

```{r data}
all_data <- get_data(location_id = c(92751, 100))
all_data$year <- lubridate::year(all_data$date_taken)
data <- all_data
```

## Locate

Metadata required about the location

```{r}
# A list of metadata for location of sampling
location <- data %>% select(
  location_id, # optional (recommended)
  location_description,
  grid_reference
)
location <- location %>% unique()
head(location)
```

## Sample

Metadata required to assess each sample. (Excluding direct observations)

```{r}
# A list of questions and responses collected through observation
sample <- data %>% select(
  parameter,
  sample_id, # optional (recommended)
  date_taken
)
sample <- sample %>% unique()
head(sample)
```

## Observe

```{r}
questions <- data %>% select(
  question,
  response,
  label # optional (usually for taxonomic/species name)
)
questions
```


## Metrics

If applicable...

```{r indices}

indices_function <- function(data) {
  require(dplyr)
  require(tidyr)
  require(magrittr)
  require(tibble)
  # Some calculated index e.g. summary of sample responses:
  # index = sum(data$response, na.rm = TRUE)
  index <- NA

  data <- tibble(
    response = NA,
    question = NA,
    sample_id = NA
  )
  return(data)
}
indexes <- indices_function(data)
indexes
```

## Predictors

```{r predictors}
predictors <- tibble(
  PREDICTOR_1 = NA,
  PREDICTOR_2 = NA,
  location_id = NA
  # ...
)
knitr::kable(predictors)
```

## Predict

```{r prediction}
prediction_function <- function(predictors) {

  # Generates prediction based on predictors...
  prediction <- tibble(
    question = NA,
    response = NA,
    location_id = NA
  )

  return(prediction)
}
predictions <- prediction_function(predictors)
predictions
```

```{r together, echo=FALSE}

# All Together Now
sample_info <- data %>%
  select(names(sample), names(location)) %>%
  unique()

if (!is.null(indexes)) {
  indexes <- right_join(sample_info, indexes,
    by = c("sample_id" = "sample_id")
  )
} else {
  indexes <- data
}

predictions <- right_join(sample_info, predictions,
  by = c("location_id" = "location_id")
)
predictions$response <- as.character(predictions$response)

combined_data <- bind_rows(questions, predictions, indexes)
```


## Assessment

```{r assessment_table}

assessment_table <- tibble(
  asesssment = c(
    "high",
    "good",
    "moderate",
    "poor",
    "bad"
  ),
  value = c(
    0.80,
    0.60,
    0.40,
    0.20,
    0
  )
  # ...
  # "red" = 0.33
  # "amber" = 0.5
  # "green" = 0.66
  # ...
  # "pass" = 0.5
  # "fail" = 0.0
)

assessment_table
```

## Assess

```{r assessment}
assessment_function <- function(data, assessment_table) {
  if (nrow(data %>% filter(question == "MY_METRIC")) == 0) {
    return(NULL)
  }

  require(dplyr)
  require(tidyr)
  require(magrittr)
  require(tibble)
  # Transform data -----------------------------------------------------------
  data <- data %>%
    select(.data$sample_id, .data$question, .data$response) %>%
    filter(.data$question %in% c("MY_PREDICTION"))
  data$response <- as.numeric(data$response)

  data <- data %>%
    distinct() %>%
    group_by(.data$sample_id) %>%
    pivot_wider(names_from = .data$question, values_from = .data$response) %>%
    ungroup()
  data[is.na(data)] <- 0

  data$eqr <- data$MY_METRIC / data$MY_PREDICTION

  class <- cut(data$eqr,
    breaks = c(1, assessment_table$value),
    labels = assessment_table$assesssment
  )

  assessments <- data.frame(
    sample_id = data$sample_id,
    class = class,
    eqr = data$eqr,
    status = data$status
  )

  assessments <- pivot_longer(assessments, -sample_id,
    names_to = "question", values_to = "response"
  )

  return(assessments)
}
assessments <- assessment_function(combined_data, assessment_table)
assessments
```

## Confidence

Confidence of assessment. 

```{r}

confidence_function <- function(data, aggregates = "sample_id") {
  confidence <- tibble(
    aggregates = NA,
    question = NA,
    response = NA
  )

  names(confidence)[1] <- aggregates[1]
}
```

## Aggregation

```{r}
aggregation_function <- function(data) {
  return(NULL)
}
```


## Checklist

```{r checklist}
check_list <- hera:::hera_test(standard = standard)
knitr::kable(check_list$standard_check)
```

## Update

```{r hera_update}
# No need to edit this code
model_dataframe <- hera::model_dataframe
model <- tibble(
  analysis_name = standard$parameter,
  standard = list(standard),
  location = list(location[1, ]),
  sample = list(sample[1, ]),
  validation_function = NA,
  indices_function = list(indices_function),
  prediction_function = list(prediction_function),
  assessment_function = list(assessment_function),
  confidence_function = list(confidence_function),
  indices = list(indexes[indexes$sample_id == indexes$sample_id[1], ]),
  assessment_table = list(assessment_table),
  questions = list(questions[1, ]),
  predictors = list(predictors[1, ]),
  predictions = list(predictions[predictions$location_id == predictions$location_id[1], ])
)

model_dataframe <- model_dataframe[model_dataframe$analysis_name != standard$parameter, ]

model_dataframe <- bind_rows(model_dataframe, model)
new_model_dataframe <- model_dataframe
new_model_dataframe
usethis::use_data(model_dataframe, overwrite = TRUE)
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=TRUE, eval=FALSE}
# No need to edit this code
launch_app(new_model_dataframe = model_dataframe, data = all_data)
```
