---
title: "Fish farm assessment"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Fish farm assessment}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

## Description

Basic details about the assessment. Update the 'response' values as required.

```{r description}

description <- tribble(
  ~question, ~response,
  "name_short", "Fish farm assessment",
  "name_long", "Fish farm assessment",
  "parameter", "Marine Benthic Inverts",
  "status", "prototype",
  "type", "compliance"
)

description
```

## Input

A list of questions required to run the assessment.

```{r input}

input <- tibble(
  parameter = c("Benthic DNA","Benthic Invertebrates", "Sediment Chemistry"),
  question = c("dna_reads","taxon_abundance", "toxicity"),
  response = c("BBB-CCC-AAA",2, 10),
  type = c("character","number", "number"),
  label =  c(NA,"bug",NA),
  sample_id = NA,
  location_id = NA,
  location_description = NA,
  grid_reference = NA,
  water_body_id = NA,
  year = NA,
  latitude = 0,
  longitude = 0,
  date_taken = NA,
  units = NA, 
  min = NA, 
  max = NA,
  source = c("sepa_ecology_results",
             "sepa_ecology_results", 
             "sepa_ecology_results")
)
input

```

## Assessment

If applicable, write a function to assess your input data and return an outcome. For example, a metric, statistic, prediction etc.

```{r assessment}

assessment_function <- function(data) {
  # Calculated some statistic...
  # Note, any non-standard base R library must be call using require().
  # require(dplyr)
  # require(tidyr)
  # require(magrittr)
  # require(tibble)
  # data <- data %>%
  #   group_by(sample_id) %>% 
  #   summarise(question = "n_count",
  #             response = n(),
  #             sample_id = unique(sample_id)
  #             )
  
  return(data)
}
```

## Outcome

The outcome of your assessment.

```{r output}
outcome <- assessment_function(input)
outcome
```

## Check

Run checks on the assessment.

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description) 
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update}
# No need to edit this code
# hera:::update_catalogue(
#   description = description,
#   input = input,
#   assessment_function = assessment_function,
#   output = outcome
# )
```

After **updating the catalogue, rebuild the package**, click on Build > Install and Restart menu or 'Install and Restart' button in the Build pane.

## Test

This section tests if this assessment is usable using `assessment` function.

```{r}
# No need to edit this code
# assess(
#   data = hera::demo_data,
#   name = description$response[description$question == "name_long"]
# )
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
