---
title: "Macroinvertebrate Metrics"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Macroinvertebrate Metrics}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

## Description

Basic details about the assessment. Update the 'response' values as required.

```{r description}

description <- tribble(
  ~question, ~response,
  "name_short", "Macroinvertebrate Metrics",
  "name_long", "Freshwater Macroinvertebrate Metrics",
  "parameter", "River Invertevates",
  "status", "prototype"
)

description
```

## Input

A list of questions required to run the assessment.

```{r input}

input <- tribble(
  ~sample_id,
  ~question,
  ~response,
  ~label,
  ~parameter,
  ~type,
  ~max,
  ~min,
  ~source,
  "12345",
  "Taxon abundance",
  "12",
  "Baetidae",
  "River Family Inverts",
  "number",
  NA,
  NA,
  "sepa_ecology_results"
)
input
```

## Assessment

If applicable, write a function to assess your input data and return an outcome. For example, a metric, statistic, prediction etc.

```{r assessment}

assessment_function <- function(data) {
  # Calculated invert metrics...
  # Note, any non-standard base R library must be call using require().
   require(dplyr)
  input <- data
  input <- dplyr::filter(input, parameter == "River Family Inverts") 
  input <- dplyr::filter(input, question == "Taxon abundance") 
  input$TAXON <- input$label 
  input <-  dplyr::rename(input,  "SAMPLE_ID" = sample_id)
  input$RESULT <- input$response
  input <-  dplyr::select(input, SAMPLE_ID,
                 TAXON, RESULT)
  input <-  dplyr::mutate_all(input, na_if, "")
  input <- input[complete.cases(input),]
  input$RESULT <- as.numeric(input$RESULT)
  epsi <- macroinvertebrateMetrics::calcEpsi(input)
  whpt <- macroinvertebrateMetrics::calcWhpt(input)
  whpt$RESULT <- as.character(whpt$RESULT)
  output <-  dplyr::bind_rows(epsi, whpt)
  output <-  dplyr::select(output, SAMPLE_ID,
                   DETERMINAND, RESULT)
  names(output) <- c("sample_id", "question", "response")
  row.names(output) <- NULL
  return(output)
}
```

## Outcome

The outcome of your assessment.

```{r output}
outcome <- assessment_function(input)
outcome
```

## Check

Run checks on the assessment.

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description) 
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = outcome
)
```

After **updating the catalogue, rebuild the package**, click on Build > Install and Restart menu or 'Install and Restart' button in the Build pane.

## Test

This section tests if this assessment is usable using `assessment` function.

```{r}
# No need to edit this code
assess(
  data = hera::demo_data,
  name = description$response[description$question == "name_long"]
)
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
