---
title: "Macroinvertebrate Metrics"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Macroinvertebrate Metrics}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

## Description

Basic details about the assessment. Update the 'response' values as required.

```{r description}
description <- tribble(
  ~question, ~response,
  "name_short", "Macroinvertebrate Metrics",
  "name_long", "Freshwater Macroinvertebrate Metrics",
  "parameter", "River Invertebrates",
  "status", "prototype",
  "type", "metric"
)

knitr::kable(description)
```

## Input

A list of questions required to run the assessment.

```{r input}
input <- tribble(
  ~sample_id,
  ~question,
  ~response,
  ~label,
  ~parameter,
  ~type,
  ~max,
  ~min,
  ~source,
  "12345",
  "Taxon abundance",
  "12",
  "Baetidae",
  "River Family Inverts",
  "number",
  NA,
  NA,
  "sepa_ecology_results",
  "12345",
  "Live abundance",
  "12",
  "Baetidae",
  "BANKSIDE_INVERTS",
  "number",
  NA,
  NA,
  "sepa_ecology_results"
)

data <- input
data <- select(data, question, response)
data <- data[!duplicated(data[, c("question")]), ]

knitr::kable(data)
```

## Assessment

If applicable, write a function to assess your input data and return an outcome. For example, a metric, statistic, prediction etc.

```{r assessment, collapse=TRUE, echo=TRUE}
assessment_function <- function(data, ...) {
  # Calculated invert metrics...
  # Note, any non-standard base R library must be call using require().
  require(dplyr)
  require(whpt)
  require(macroinvertebrateMetrics)
  input <- data
  input$label <- trimws(input$label)
  input <- dplyr::filter(
    input,
    parameter %in% c(
      "River Family Inverts",
      "BANKSIDE_INVERTS"
    )
  )
  input <- dplyr::filter(input, question %in% c(
    "Taxon abundance",
    "Live abundance"
  ))
   if(nrow(input) < 1) {
     return(NULL)
   }
  input <- ungroup(input)
  input <- dplyr::select(
    input, "sample_id", "question", "response", "label", "parameter"
  )

  output <- macroinvertebrateMetrics::calc_metric(input)

  # riverfly <- macroinvertebrateMetrics::calc_riverfly(input)
  
  # output <- dplyr::bind_rows(whpt, riverfly)

  output <- dplyr::select(
    output, "sample_id", "question", "response","parameter"

  )
  names(output) <- c("sample_id", "question", "response","parameter")
  return(output)
}
```

## Outcome

The outcome of your assessment.

```{r output}
# data <- get_data(location_id = 8175)
# outcome <- assessment_function(data)
outcome <- assessment_function(input)
outcome_table <- select(outcome, question, response)
outcome_table <- outcome_table[!duplicated(outcome_table[, c("question")]), ]
knitr::kable(outcome_table)
```

## Check

Run checks on the assessment.

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description)
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = outcome
)
```

After **updating the catalogue, rebuild the package**, click on Build > Install and Restart menu or 'Install and Restart' button in the Build pane.

## Test

This section tests if this assessment is usable using `assessment` function.

```{r, results='hide'}
# No need to edit this code
# assess(
#   data = hera::demo_data,
#   name = description$response[description$question == "name_short"]
# )
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
