---
title: "Fish farm assessment"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Fish farm assessment}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = FALSE
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
#  library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

<!-- Create a new regulatory assessment by updating the code info below, for guidance see https://ecodata1.github.io/hera/articles/development_guide.html  -->

## Description

Basic details about the assessment.

```{r description}
description <- tribble(
  ~question, ~response,
  "name_short", "MPFF Compliance",
  "name_long", "Marine Pen Fish farm Compliance",
  "parameter", "Marine Benthic Inverts",
  "status", "prototype",
  "type", "compliance"
)

knitr::kable(description)
```

## Input

A list of questions required to run the assessment.

```{r input}
file <- system.file("extdat",
  "demo-data/220421-SelfMon-N4952-CAV1-Enhanced.xlsx",
  package = "hera"
)
input <- survey_import(file)

# Only display columns vital to run example (min, max, source, type are not
# required)
input_display <- select(
  input,
  project_id,
  sample_id,
  question,
  response,
  label,
  parameter
)
# Only display one example of each question
input_display <- input_display[!duplicated(input_display[, c("question")]), ]
knitr::kable(input_display)
```

## Assessment

Function code used to run this assessment.

```{r assessment, message=FALSE, warning=FALSE, collapse=TRUE, echo=TRUE}
assessment_function <- function(data, ...) {
  args <- list(...)
  if (length(args) > 0) {
    results <- kraken::kraken(data, ...)
  } else {
    results <- kraken::kraken(data, hera_format = TRUE)
  }
  return(results)
}
```

## Outcome

The outcome of your assessment.

```{r output, message=FALSE, warning=FALSE}
outcome <- assessment_function(
  input,
  hera_format = TRUE, 
  niter = 10, 
  loess = TRUE)
outcome_table <- select(outcome, question, response)
outcome_table <- outcome_table[!duplicated(outcome_table[, c("question")]), ]
knitr::kable(outcome_table)
```

## Check

Run checks on the assessment.

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description)
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

Update the catalogue of assessments to make them available.

```{r hera_update}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = outcome
)
```

After **updating the catalogue, rebuild the package**, click on Build > Install and Restart menu or 'Install and Restart' button in the Build pane.

## Test

This section tests if this assessment is usable using `assessment` function.

```{r}
# No need to edit this code
# assess(
#   data = hera::demo_data,
#   name = description$response[description$question == "name_long"]
# )
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
