---
title: "Trophic Diatom Index"
output: rmarkdown::html_vignette
editor_options: 
  chunk_output_type: console
vignette: >
  %\VignetteIndexEntry{Trophic Diatom Index}
  %\VignetteEncoding{UTF-8}  
  %\VignetteEngine{knitr::rmarkdown}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = F
)
library(hera)
library(dplyr)
library(purrr)
library(tidyr)
library(tibble)
library(magrittr)
library(testthat)
library(visNetwork)
```

## Welcome

This document has been created following the generic [assessment guidance](https://ecodata1.github.io/hera/articles/development_guide.html).

## Description

```{r setup}
description <- tribble(
  ~question, ~response,
  "name_short", "TDI",
  "name_long", "Trophic Diatom Index",
  "parameter", "River Diatoms",
  "status", "prototype"
)

description
```

## Input

Input data required for assessment. Can be from test data, real data or a short example.

```{r input}
input <- tribble(
  ~sample_id,
  ~question,
  ~response,
  ~label,
  ~parameter,
  ~type,
  ~max,
  ~min,
  ~source,
  "12345",
  "Taxon abundance",
  "12",
  "Gomphonema olivaceum",
  "River Diatoms",
  "number",
  NA,
  NA,
  "sepa_ecology_results"
)
input
```

## Assessment

Calculate the Trophic Diatom Score (TDI) for a given sample. This score is used to assess the nutrient status of freshwaters.

```{r assessment, echo=TRUE}
assessment_function <- function(data) {
  require(dplyr)
  require(tidyr)
  require(magrittr)
  require(tibble)

  darleq3_taxa <- darleq3::darleq3_taxa
  output <- filter(data, question == "Taxon abundance")
  output <- inner_join(output, darleq3_taxa, by = c("label" = "TaxonName"))

  output <- output %>%
    group_by(sample_id) %>%
    mutate("total_count" = sum(as.numeric(response)))

  # Group data to get TDI scores for each taxa
  output <- output %>%
    group_by(sample_id, label, total_count, Saline) %>%
    summarise("TDI" = (total_count / 100 * as.numeric(response)) * TDI4,
               .groups = "drop")

  # Mean TDI scores for each taxa and work out percentage for other
  # metrics
  output <- output %>%
    group_by(sample_id, total_count) %>%
    summarise(
      question = c(
        "TDI4",
        "Saline"
      ),
      "response" = c(
        mean(TDI, na.rm = T),
        unique(total_count) / 100 * sum(Saline)
      ),
      .groups = "drop")
  output <- ungroup(output)
  output <- output %>% select(-total_count)

  return(output)
}
```

## Output

```{r output}
output <- assessment_function(input)
output
```

## Check

Run checks on the assessment:

```{r checklist}
# No need to edit this code
# Format description
standard_format <- hera:::hera_format(description = description) 
# Check description
check_list <- hera:::hera_test(description = description)
knitr::kable(check_list$standard_check)
```

## Update

```{r hera_update}
# No need to edit this code
hera:::update_catalogue(
  description = description,
  input = input,
  assessment_function = assessment_function,
  output = output
)
```

## Hera

This section tests if this assessment is usable using `hera` function

```{r}
hera(hera::demo_data)
assessment(data = hera::demo_data, name = "Trophic Diatom Index")
```

## Launch app

Below is an interactive application displaying the results of your assessment.

```{r launch_app, echo=FALSE, eval=FALSE}
# No need to edit this code
# launch_app(new_catalogue = catalogue, data = data)
```
